---
title: "Stat 3540/5440: Statistical Methods for Study Design and Analysis Syllabus"
subtitle: "Fall 2026"
---

```{r}
#| include: false
#| eval: true
library(tidyverse)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# Format: YYYYMMDD

# What are the full dates of the semester?
semester_dates <- seq(ymd(20260824), ymd(20261218), by = 1)

# Weekday(s) of class
class_wdays <- c("Tue", "Thu")

# Breaks
not_here_dates <- c(
  # Break
  seq(ymd(20261121), ymd(20261130), by = 1),
  ymd(20260907),
  ymd(20261111)
  )

# Dates of Exams
exam_dates <- c(ymd(20260917), #midterm1
                ymd(20261020), #midterm2
                ymd(20261215), #final
                ymd(20261217) #final
                )
finals_week <- seq(ymd(20261213), ymd(20261218), by = 1)
project_dates <- ymd(20261210)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
}

# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20260801), ymd(20261231), by = 1))  |> 
  mutate(mon      = lubridate::month(date, label = T, abbr = F),
         wkdy     = weekdays(date, abbreviate = T),
         wkdy     = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri", "Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         project  = date %in% project_dates,
         exams    = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk  = date %in% finals_week,
         day      = lubridate::mday(date), 
         week     = wom(date)
         ) |> 
  mutate(category = case_when(project ~ "Project",
                              exams ~ "Exam",
                              not_here ~ "Cal Poly Holiday",
                              wkdy %in% class_wdays & !not_here & !exam_wk & date %in% semester_dates ~ "Class Day",
                           #   date %in% ymd(20250121) ~ "Class Day",
                              semester ~ "Finals",
                              TRUE ~ "NA")
         )


class_cal <- Cal |> 
  # mutate(week = case_when(mon == "March"~ 1,
  #                         .default = week),
  #        mon = replace_values(mon, "March" ~ "April")
  #        ) |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0.8, 
            aes(fill = category), 
            color="black", 
            linewidth = 0.45
            ) +
  facet_wrap(~ mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks = NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="steelblue", 
                             "Finals"="white",
                             "Cal Poly Holiday" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Exam"="orange3",
                             "Project"="#c994c7"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "Cal Poly Holiday", "Class Day","Exam","Project"))
  
# class_cal

exam_days <- Cal |> 
  filter(category == "Exam") |> 
  mutate(topic = c("Midterm Exam I", 
                   "Midterm Exam II",
                   "Sec 03 Final Exam 4-6:30pm",
                   "Sec 02 Final Exam 10am-12:30pm \\& Sec 01 Final Exam 7-9:30pm"
                   )
         ) |> 
  rowwise() |> 
  mutate(sem_week = isoweek(date),
         Date = paste(format(date, "%b %e"), sep = ",", collapse = ", "))

project_days <- Cal |> 
  filter(category == "Project") |> 
  mutate(topic = c("Project Presentations")
         ) |> 
  rowwise() |> 
  mutate(sem_week = isoweek(date),
         Date = paste(format(date, "%b %e"), sep = ",", collapse = ", "))



class_days <- Cal |> 
  filter(date %in% semester_dates,
         category %in% c("Class Day")
         ) |>  
  mutate(sem_week = isoweek(date)) |> 
  group_by(sem_week) |> 
  summarize(Date = paste(format(date, "%b %e"), sep = ",", collapse = ", ")) |> 
  ungroup() |>
  mutate(topic = c(
    "Introduction to Design of Experiments",
    "Completely Randomized Design (CRD)",
    "Power and Sample Size + Factorials",
    "Factorials",
    "More Complex Experimental Designs",
    "More Complex Experimental Models",
    "Extensions and Considerations",
    "Extensions and Considerations",
    "Observational Studies",
    "Intro to Surveys and Writing Survey Instrucments",
    "Writing Survey Instruments",
    "Survey Sampling",
    "Survey Sampling and Survey Weights",
    "Putting it all Together!",
    "Putting it all Together!")) |> 
  bind_rows(project_days) |> 
  bind_rows(exam_days) |> 
  arrange(sem_week, Date) |> 
  select(Date, Topic = topic)


# Weekday(s) of office hours
officehr_wdays <- c("Mon", "Wed")

# Create a data frame of dates, assign to Cal
Cal2 <- tibble(date = seq(ymd(20240107), ymd(20240113), by = 1))  |> 
  mutate(mon      = lubridate::month(date, label = T, abbr = F),
         wkdy     = weekdays(date, abbreviate = T),
         wkdy     = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri", "Sat"), # make sure Sunday comes first
         week     = wom(date),
         classday = wkdy %in% class_wdays,
         officehrday = wkdy %in% officehr_wdays,
         officehrtime = c(NA, "1-3pm", NA, "9-11am", NA, NA, NA)
  )


classday_cal <- Cal2 |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 0.15, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0.8, 
            aes(fill = classday), 
            color="black", 
            linewidth = 0.45,
            show.legend = F
  ) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values=c("white", "steelblue")
                    )
# classday_cal

officehr_cal <- Cal2 |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 0.15, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0., 
            aes(fill = officehrday), 
            color="black", 
            linewidth = 0.45,
            show.legend = F
  ) +
  geom_text(aes(label = officehrtime), color = "black") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values=c("white", "gray")
  )
# officehr_cal

```

## Communication

**Dr. Charlotte Zilber Mann**^[Why do I include my middle name? It's my mom's last name!] (she/her/ella)

You can call me Dr. C, Professor/Prof. C., Dr. Mann, or Professor.^[As an FYI for folks who don't know, it is considered rude to use "Ms." or "Mrs." when addressing instructors. I will get annoyed if you do this and will correct you.]


`r fontawesome::fa("envelope")` Email:
[czmann\@calpoly.edu](mailto:czmann@calpoly.edu?subject=Stat%20331/531)

`r fontawesome::fa("building")` Office: Building 25 Office 201

**Email:**

-   Only email me with your @calpoly.edu email – I cannot respond to
    other email addresses
-   Please **start your email subject line with “\[STAT 3540\]”** or **“\[STAT 5440\]”** to get my attention!
-   I will do my best to respond within 24 hours during the week
-   If you send me an email past 6pm or on the weekend, I will likely
    not respond until the morning of the next working day
- If I don't respond within 48 hours, please email me again!
-   *You should only email me about things that relate to you as an
    individual.* Any other questions should be posted on the course
    Discord (more details below).

**Course Discord Page:**

For questions of general interest, such as course clarifications or
conceptual questions, please use the course [Discord page](discord-instructions.html) (you will join
this Week 1). If you have a question, someone else in class does too! I encourage you all to respond to each other. I will do my best to check Discord twice each weekday during working hours (8-6) to ensure that questions are being answered accurately.

I encourage you to give your post a concise and informative initial
sentence, so that other people can find it. You may post snippets of code, errors or problem attempts on Discord, but please do not post full solutions.

`r fontawesome::fa("brain")` While your posts are not anonymous, in this case there is no such thing as a bad question!

## Course Logistics

**Class Meeting Times:** Tuesdays / Thursdays 

```{r}
#| eval: false
#| include: false
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 1.5
#| out-width: 75%
#| fig-align: center
#| fig-pos: H
classday_cal
```

-   Section 01: 8:00am - 9:50am
-   Section 02: 10:00am - 11:50am
-   Section 03: 2:00pm - 3:50pm

**Classrooms**:

- Sections 01 & 02: 180-272
- Section 03: 180-0102 

**Office Hours** are held in my office (25-201) during the following
times:

<!-- ::: column-margin -->

<!-- {{< video https://www.youtube.com/embed/yQq1-_ujXrM?si=sFpcpfT9Cey2-Eaf >}} -->

<!-- Introducing Faculty Office Hours -->

<!-- ::: -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 1.5
#| out-width: 90%
#| fig-align: center
#| fig-pos: H
officehr_cal
```

<!-- | Day        | Time                | -->

<!-- |------------|---------------------| -->

<!-- | Mondays    | 3:10pm - 4:00pm     | -->

<!-- | Wednesdays | 1:10pm - 2:00pm     | -->

<!-- | Thursdays  | 2:30pm - 3:30pm   | -->


::: column-margin
https://lsc.cornell.edu/consult-with/how-to-use-office-hours/

:::

- **Office hours are for you!** During this time, I am available to talk about anything related to the course. 
- **Office hours are also helpful for me…** this is my chance to get to know you all better and also get a good idea of what is most challenging and most interesting to you.
- See [this resource](https://lsc.cornell.edu/consult-with/how-to-use-office-hours/) for ways that students use office hours. 
- One thing I will not do during office hours is tell you whether an answer to a problem is “correct” or not. We can talk through your thinking that led to your solution, but will leave the final assessment to once you have submitted an assignment.
- **Note that office hours are not just for when you have content questions.**
Stop by to introduce yourself, ask questions about the broader field of
statistics, or share what you are working on!
- I am always open to students joining remotely via Zoom -- just email me.
- I *LOVE* office hours! Please come by -- you are never inconveniencing me.

## Course Description

Study designs to address statistical research questions. Ethical considerations when designing
studies and analyzing data. Experimental designs, completely randomized, block designs, factorial,
and nested designs. Survey planning, execution, and analysis. Survey research, non-sampling,
and sampling errors. Use of weights in surveys. Estimate parameters and calculate sample size.
Course may be offered in classroom-based, online, or hybrid format

**Prerequisites**

STAT 252, STAT 1220, STAT 302, STAT 313 or STAT 3520.

## Learning Objectives

By the end of this course, you should be able to:

- Communicate the ethical implications of research questions, study designs, and analysis choices.
- Determine which type of study (observational or experimental) is appropriate given a research context.
- Design and explain a plan for conducting an experiment (e.g., RCBD) or a sample survey (e.g., stratified sample design).
- Analyze data by applying an appropriate model that accounts for the study design and attempts to isolate and quantify the effects of the factors of interest.
- Describe the limitations of study designs and propose alternative designs.
- Produce a written and/or oral summary of an appropriate study design and analysis.

## Course Materials

**Course Webpage (Canvas)**

All course material will be made available on our Canvas page. Canvas will
also contain detailed weekly schedules throughout the term. You are responsible for
checking the Canvas page every day and making sure you see Canvas Announcements.

I will also be posting material on [this website](https://manncz.github.io/stat-3540-calpoly/) (which is actually where you are right now!) for easy access to materials. Please note that **this website does not include deadlines** and links to all material will additionally be posted to Canvas. Canvas will always contain the most updated information about course deadlines and details, so always defer to these. 

**Class Preparation**

For each topic, you will prepare for class by `r fontawesome::fa("video")` watching 1-3 short lecture videos following a set of *lecture slide notes* `r fontawesome::fa("book")` containing definitions and examples. You are expected to either print `r fontawesome::fa("print")` or save `r fontawesome::fa("save")` the lecture slide notes to your device and follow along, filling `r fontawesome::fa("pencil")` out your set of notes while watching the videos.

During class, I will highlight any key points from the lecture videos and we will complete activities meant to reinforce the ideas learned in the videos.

Supplemental reading from the following *open source* books will be recommended throughout the course:

- `r fontawesome::fa("book")` [A First Course in Design and Analysis of Experiments](http://users.stat.umn.edu/%7Egary/book/fcdae.pdf) by Gary W. Oehlert
- `r fontawesome::fa("book")` [Internet, Phone, Mail, and Mixed‐Mode Surveys: The Tailored Design Method](https://onlinelibrary.wiley.com/doi/book/10.1002/9781394260645) by Don A Dillman, Jolene D Smyth, and Leah Melani Christian

**Computing**

`r fontawesome::fa("computer")` You will need access to a laptop for this course. If this requirement is limiting for you, please
contact me ASAP.

We will be using the `R` statistical software and `RStudio` IDE
throughout this course . Both `R` and `RStudio` are freely available. Download instructions are available on Canvas. If you can, please work on downloading this software before class begins.

`r fontawesome::fa("phone")` We will also be using Gradescope -- an app on your phone or computer browser.

::: callout-warning
Chromebooks and iPads will **not** be
sufficient to use `R`. If this requirement is limiting for you, please
contact me ASAP.
:::


\newpage

## Class Schedule & Topic Outline

This schedule is tentative and subject to change. See Canvas and the course [Schedule](../index.qmd) for up-to-date information. 

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-pos: 'h'
#| fig-cap: "Note that all Cal Poly holidays other than Fall Break fall on Mondays or Wednesdays, so we will have regularly scheduled class throughout the entire semester."
class_cal
```

```{r schedule}
#| echo: false
#| eval: true
#| fig-pos: 'h'
class_days |> 
kableExtra::kable(caption = "Tentative schedule of class topics and important due dates", format = "simple")
```


## Assessments

### General Evaluation Criteria

In every assignment, discussion, and written component of this class,
you are expected to demonstrate that you are intellectually engaging
with the material. I will evaluate you based on this engagement, which
means that technically correct but low effort answers which do not
demonstrate engagement or understanding will receive no credit.

When you answer questions in this class, your goal is to show that you
either understand the material or are actively engaging with it. If you
did not achieve this goal, then your answer is incomplete, regardless of
whether or not it is technically correct. This is not to encourage you
to add unnecessary complexity to your answer - simple, elegant solutions
are always preferable to unwieldly, complex solutions that accomplish
the same task.

While this is not an English class, grammar and spelling are important,
as is your ability to communicate technical information in writing; both
of these criteria will be used in addition to assignment-specific
rubrics to evaluate your work.

### Formative Assessments

#### Class Preparation, Participation, \& Professionalism

This course uses a flipped classroom model. Core content is introduced before class through assigned videos or readings, and class time is devoted to activities and practice. Coming to class prepared is essential for your own learning and for your group’s success. Preparation will occasionally be checked using short, closed-note *learning quizzes* or *entry/exit tickets*

Participation credit is earned by being present, prepared, and engaged during class. You must be in class to receive participation credit. Part of this grade also reflects professionalism, including arriving on time, being mentally present, treating others with respect, and contributing constructively to group work.

+ **2** participation grades will be dropped; think of these as "sick days".


::: callout-note
If you are feeling ill, please *do not come to class*. If you miss more than 2 class days for illness, we will work something out!

See @sec-attendance for more information. You do not need to notify me, but instead do your best to complete the in-class activity found on Canvas and get any questions answered from your classmates, Discord, and office hours.
:::

#### In-class Activities

Most class meetings will include in-class activities designed to reinforce and apply course content. Class time will be dedicated to working on these activities. You will collaborate with your group members; however, each student must submit their own original work.

+ In-class activities will be submitted and graded through Gradescope, which is integrated with Canvas. 
+ Unless otherwise noted, activities are due at **11:59pm the day after class.**

::: callout-note
Cal Poly provides access to [Gradescope](https://www.gradescope.com/). There will be more instructions on using Gradescope as a student during the first week of class.
:::

#### Lab Reports

You will complete 4-6 lab assignments (I am still figuring this out!) in groups of 3-4. Each lab will involve designing a study and collecting data, often using [The Islands](https://islands.smp.uq.edu.au/). A written report summarizing your methods and findings will be submitted for each lab. More details will be provided in class.

+ One report per group should be submitted as a PDF on Canvas.
+ See the course schedule on Canvas for tentative deadlines.


### Evaluative Assessments

#### Final Project

You will complete a paired final project in which you design and propose a study, including all necessary details. The project will culminate in a 10-minute presentation during the last week of class. Additional details and expectations will be provided later in the term.

#### Exams

There will be **two midterm exams** and a **final exam**, designed to assess your individual understanding and progress in the course. You may bring a one-sided notesheet to each midterm exams and a two-sided notesheet to the final exam. Exam dates are listed in the course calendar.


:::callout-note

If you have a known conflict with an exam, please discuss it with me **at least three weeks prior to the exam date**. (See @sec-examconflict for more information.)
:::

### Course Grade

Your grade in STAT 3540/5440 will contain the following components:

```{r}
#| echo: false
#| fig-loc: "H"

tribble(
  ~Assignments,  ~Weight,
  "Class Preparation, Participation, & Professionalism", "5%",
  "In-class Activities", "10%",
  "Lab Reports", "15%",
  "Final Project", "10%",
  "Midterm Exams I & II *", "35%",
  "Final Exam", "25%"
) |> 
  knitr::kable() |> 
  kableExtra::kable_styling(full_width = F)
```

\* At the end of the semester, your higher-scoring midterm exam will be weighted to 20% of your final grade, while the lower-scoring midterm will be weighted to 15%.


Lower bounds for grade cutoffs are shown in the following table. It is possible that I could lower any of the grade cutoffs. You will only have the opportunity for grade rounding to the nearest tenth of a percent if you completed all assignments in the course. 

| Letter grade | X +  | X   | X - |
|--------------|------|-----|-----|
| A            | .    | 93  | 90  |
| B            | 87   | 83  | 80  |
| C            | 77   | 73  | 70  |
| D            | 67   | 63  | 60  |
| F            | \<60 |     |     |

Interpretation of this table:

-   A grade of 85 will receive a B.
-   A grade of 77 will receive a C+.
-   A grade of 70 will receive a C-.
-   Anything below a 60 will receive an F.

## Course Policies

### Attendance & Participation {#sec-attendance}

As stated above, **if you are feeling ill, please do not come to class**. Two participation grades will be dropped as "sick days." If you *really* just have a cold and *really* want to come to class be polite to your classmates and wear a mask!

Here’s what you should do if you do miss a class:

-  You do not need to notify me, unless there is a special circumstance you want to discuss.
-   Complete the in-class activities on your own, which are posted on Canvas.
- Check Canvas for any other necessary notes or changes to assignments.
-   Talk to a classmate to figure out what information you missed.
-   Post on Discord with any questions you have after reviewing notes and activities.
- Come by office hours when you are feeling better.

:::callout-tip
## Don't be a stranger!

If you miss a bunch of classes, please do come talk to me. I’m working from the assumption that you care and are trying, but something is getting in
your way (health issues? college stress? work?); let’s figure out what
that is and how I can help.
:::

### Accessibility and Accommodations

I have taken numerous steps to ensure that all of the materials presented in this course are accessible to all participants, regardless of physical or learning disabilities. I know that everyone is unique and I may have unintentionally overlooked something that limits access to some materials or activities. Please let me know if you cannot access any content. If you find you need additional accommodations to complete the required course work, please contact me as soon as possible! I want to support your success. 

If you wish to request disability-related accommodations such as extra exam time for this or any other course, please contact the [Disability Resource Center](https://drc.calpoly.edu/). It's important to do this with as much advanced notice as possible, so you have full access to your course materials and activities in a timely manner. **Please also email me** to let me know that I should expect an email from the DRC to get you the accommodations that you need.


### Inclusive Classroom 
Because data are collected by and about humans, data necessarily
encode aspects of our proclivities and biases. As a result, this course may touch upon
difficult topics related to race, gender, inequality, class, and oppression. We each come into
this class with different perspectives that can be shared to enhance our understanding of
these issues. I ask that you enter these conversations with respect, curiosity, and cultural
humility. You should be open to alternative perspectives and be willing to revise beliefs that
are based on misinformation. As a general rule, your ideas and experiences can always be
shared during these conversations, but please refrain from dismissing the experiences of
others. Personal attacks of any kind will not be tolerated.

### Group Work

This course will rely heavily on in-class group work. You will be expected to work with classmates on in-class material. If you have concerns about random group assignment, please
talk to me at the start of the term.

The community standards for student-to-student and student-to-instructor interactions include the following:

- Listen actively and attentively.
- Ask for clarification if you are confused.
- Challenge one another respectfully.
- Gracefully accept constructive criticism.
- Take responsibility for the quality of the discussion and the work.
- Do not monopolize discussion.
- Acknowledge that everyone has something to offer.
- Speak from your own experience, without generalizing.


### Late Policy


**Everyone will be given a “bank” of 4 *deadline extensions*** which can be used on in-class activities or lab reports at any point in the quarter. **One deadline extension will get you a 24 hour extension on a given assignment.** You can use all of your deadline extensions on one assignment or split them up in any way. 


-	Like how the “real world” works, you must request a deadline extension **before the deadline.** To use your deadline extensions, you need to email me **before the assignment deadline.** You don't need to get a response from me, you just need to email me at least a minute before the deadline. 
-	If you do not email me before the assignment deadline (or you run out of late days), you will be docked 20% for each day that the assignment is late. (e.g. you can get a maximum of 80% on the assignment if it is submitted within 24 hours after the original due date and time.)
- In order to use deadline extensions on group *lab reports*:
    1. **everyone** in your lab group must agree to use one of their extensions. 
    2. You **must cc everyone** in your group to your email to me asking for the extension.

A note on this policy: my priority is for all of you to engage with and
understand the course materials. The best way to achieve this is to
complete all assignments and receive feedback on your work. At the same
time, I understand that things happen outside of our control and that
you are balancing many responsibilities.

::: callout-warning
### Automatic Canvas Settings

Canvas is set up to automatically input 0% for missing assignments (as
an incentive to go complete the assignment) and apply the 20% grade
deduction policy. I will need to manually adjust your grade when you use
your "deadline extensions." You are responsible for double checking your
grade to make sure I didn't miss anything.
:::

If you find yourself with extenuating circumstances beyond the defined
late policy, please email me before the due date.


### Exam Conflicts {#sec-examconflict}

If you have a known conflict with an exam, please discuss it with me **at least three weeks prior to the exam date**. If an illness or a family emergency arises, please
let me know as soon as possible, and we will work out a solution. This may involve taking the
exam at a later date or replacing the missed exam score with the final exam score. If you are
unable to take the final exam, you will likely receive an Incomplete in the course.

## Course Expectations


**For Students, I expect:**

- You will ask lots of questions.
- You will do the reading, video watching, and other assignments outside of class and be prepared for each class meeting.
- You will engage in some productive struggle/challenge when working on assignments (i.e. not give up and google or use ChatGPT immediately!)
- If you find yourself struggling in general or feel that you are falling behind, please reach out to me as soon as possible so that we can develop a plan for your success together. 
- You will take advantage of the resources that I provide and to seek out additional resources if you find that you are struggling with the material. You do not need to seek out additional resources all on your own – we can figure out what additional resources you may need.  
- You will be doing your best to engage with what we are all doing together during class. This means that you are using technology (i.e. laptops/tablets/phones) for class activities and not in ways that distract you. 
- You will approach the course with academic integrity.
- You will come to class on time and work on class material with your peers.
- Everyone can learn the material in this course.

**For Me:**

- I am responsible for ensuring that all students feel welcome and valued in the classroom.
- I am responsible for providing the resources necessary so that all students in the course can achieve the learning objectives.
- I am responsible for providing prompt and clear feedback on your coursework.
- I welcome and appreciate any actionable feedback regarding how the class is meeting your learning needs. We can discuss ways to adjust the course to ensure it is supporting your learning and do more of anything that is working well!



## Academic Integrity and Class Conduct

**Our academic community is at its best when we treat each other with fairness, honesty, respect, and trust. Unfortunately, sometimes students slip up and do something that gives themselves (or someone else) an unfair advantage over other students.  Such actions will not be tolerated in this class.**

It is most likely that a student will slip up if they are feeling overwhelmed or unsure of how to approach a problem that they are stuck on. So **let's be proactive to prevent those situations!!** If you are helpless in an assignment or in studying, post on discord, come to office hours, or email me or one of your classmates! I am working hard to provide all of the resources you need to succeed


However, **the college doesn't really care if you break the rules on purpose or by accident.** If I suspect you have done something that violates academic integrity on any graded elements of the course (included those graded for completion), you will receive an email to tell you about my suspicion and how the situation will be handled. **Typically, students earn a 0 for the  assignment on which academic misconduct is found.** University policy dictates that we must report every instance of suspected academic dishonesty to the Office of Student Rights and Responsibilities, no matter how small. For more
information on academic misconduct and what constitutes cheating and
plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

It is important to note that paraphrasing or quoting another's work
without proper citation is a form of academic misconduct. This includes
the `R` code produced by someone else! Writing code is like writing a
paper, it is obvious if you copied-and-pasted a sentence from someone
else into your paper because the way each person writes is different.

Even inadvertent or unintentional misuse or appropriation of another’s
work (such as relying heavily on source material that is not expressly
acknowledged) is considered plagiarism. This includes using **Generative AI, which should only be used to help you problem solve just as a tutor or peer would, and not as a substitute for your own work.** If you are
struggling with an assignment, please reach out
to me and post to Discord. If you use Generative AI to help, make sure that you **understand** the solution it provides.



::: callout-tip
## Always Cite Outside Resources!

In this class the assumed knowledge is the course materials, including the course textbook, coursework pages, and course slides.  If you use outside resources on assignments I expect that you will *cite those resources.* This means that

- If you use **Google** you are expected to “inform” me of any resources you used by pasting the link to the resource at the end of your assignment with a note of how it was used.
- If you use **AI** you are expected to “inform” me by stating that you used at the end of your assignment including **the full prompt that you gave AI (like ChatGPT)**. See an example of an appropriate citation below.
- If you **work with classmates** on an assignment, you are expected to “inform” me by including a note with their names. 
:::

**If you have any questions about using and citing sources, you are
expected to ask for clarification.**

### Generative AI Policy

Generative AI (like Chat GPT) is an extremely and increasingly powerful tool for many many tasks. It is true that in some (how many, I am not sure!) workplaces, employees are encouraged to use AI for a number of tasks. However, in this class, you are students, not employees! Your primary goal here is to learn, so I want us to think critically about how to use Generative AI (if you choose to use it at all) to **assist your learning**, not to undermine it.

:::callout-important
# No copy-paste, seriously
You **may not copy and paste any course materials into generative AI**. That includes slides and assignments. All course materials are my or other faculty's intellectual property, and we have not consented to you contributing it to generative AI model data.  
:::

#### In Class:
Class time is structured around active learning and I expect you to actively engage with the software, so **you may not use generative AI for in-class work and activities**. You have each other and me to grapple with the material during class.  If you use generative AI in class, you will not get participation credit for that day.

#### Outside of Class:

You can use generative AI outside of class on assignments. However, I encourage you to give all problems a full effort on your own, referencing the provided course materials **before** turning to AI. You will learn SO much more this way!! You should also give thought to how you are using AI, and approach it as a tutor, asking it specific questions to help you solve a problem. Again, you may not copy and paste course materials, including problems, into AI. When you do use AI to help with a problem, make sure you understand **why** the code it provides works or why the solution make sense. **If you choose to use AI, you must cite your use including the prompt that you gave.**

:::callout-tip
## Example AI citation

- 



:::


Note that work created by AI tools may not be considered original work and instead, considered automated plagiarism. It is derived from previously created texts from other sources that the models were trained on, yet doesn't cite sources. Thus, there is a high risk that you can unintentionally plagiarize other sources from the web.



#### My 2-cents

A final personal opinion is that I find YOU so much more interesting than AI! I would so much rather see and read what you are thinking and how you approach a problem than something you may think looks more “clean” from AI. 







## Learning Environment and Support

My priority is your success in this course. I understand that things happen unexpectedly and obtaining official accommodations can be slow. 

If you feel that any circumstances or factors beyond your control may affect your presence and work in this class, **please reach out to me** and we will work to accommodate your needs. You do not need to share personal information with me when you reach out, unless you would like to.


### Mental Health

National surveys of college students have consistently found that stress, sleep problems, anxiety, depression, interpersonal concerns, death of a significant other and alcohol use are among the top ten health impediments to academic performance. If you are experiencing any mental health issues, I and Cal Poly are here to help you. Cal Poly’s Counseling Services (805-756-2511) is a free and confidential resource for assistance, support and advocacy [https://chw.calpoly.edu/counseling](https://chw.calpoly.edu/counseling).

### Support Services

If you are experiencing food insecurity, housing instability, or other
challenges that may impact your ability to succeed in this course, please refer
to the resources listed on Canvas under ["Student Support and Basic Needs"](https://canvas.calpoly.edu/courses/191796/pages/student-support-and-basic-needs).
These resources provide a range of essential support services, including
emergency financial assistance, counseling, and academic support.