---
title: "Activity 2.1: ANOVA & Statistical Model for a CRD"
subtitle: "Traffic Signals"
execute: 
  error: true
  message: false
  warning: false

author: Your Name
echo: true
embed-resources: true
format: typst
---
::: callout-note
Submit via Gradescope, don't forget to assign pages to the questions in the outline!

Note: while you are using the .qmd or Word template the links won't work -- make sure to also reference the online instructions!!
:::

*I worked on this activity with:* ______________________________

A traffic engineering study was conducted to investigate traffic delay at urban intersections. Three types of traffic signals were studied: pretimed, semi-actuated, and fully actuated. Fifteen intersections were completely randomly assigned to one of the three types of signals. The response variable was the average stopped time per vehicle (seconds) at each intersection.

```{r}
#| label: setup

# load the packages we need for this activity
library(tidyverse)
library(emmeans)
library(knitr)
library(janitor)
```

This code below reads in the data (`read_csv()`), makes the variable names nicer in R (`clean_names()`), and makes sure sure the order of the levels
of the treatment are in a nice order for us (`factor()`).

:::{.callout-note}
`factor` is a special data type in R. This is distinct from our term "factor" in relation to study design.
:::

```{r}
traffic_data <- read_csv("PATH/02_traffic_data.csv") |> 
  clean_names() |> 
  mutate(traffic_signal = factor(traffic_signal, 
                                 levels = c("Pretimed", 
                                            "Semi-actuated", 
                                            "Fully actuated")))
```

## a. Study Structure

Describe the following components of the study.

-   Treatment structure

\vspace{0.6in}

-   Design structure

\vspace{0.6in}

-   Response variable

\vspace{0.4in}

## b. Complete the skeleton ANOVA table below

Make sure to use context of the data!

| Source of Variation (SV) | DF: Total = \_\_\_\_\_\_\_\_\_\_\_ |
|--------------------------|------------------------------------|
|                          |                                    |
|                          |                                    |
|                          |                                    |
|                          |                                    |
|                          |                                    |

## c. ANOVA Table

```{r}
# chunk for calculations if desired
```

Use the data provided to calculate the following (you can use R):

$\bar y_{\cdot\cdot}$ = \_\_\_\_\_\_\_\_\_

$\bar y_{1\cdot}$ = \_\_\_\_\_\_\_\_\_

$\bar y_{2\cdot}$ = \_\_\_\_\_\_\_\_\_

$\bar y_{3\cdot}$ = \_\_\_\_\_\_\_\_\_

for $i=1$ (Pretimed), $i=2$ (Semi-actuated), $i=3$ (Fully actuated)

-   Show (by hand, using a calculator) how the SSE is calculated:

```{r}
# se calculation here if you want
```

\vspace{0.6in}

Then, use your SSE calculation and the provided information to fill out the ANOVA table below. *You can check your answers by running the ANOVA in R, but know where the values come from.*

| Source | DF  | SS      | MS  | F   | Prob \> F |
|--------|-----|---------|-----|-----|-----------|
|        |     |         |     |     | \<0.0001  |
|        |     |         |     |     |           |
| Total  | 14  | 1340.45 |     |     |           |

## d. Statistical Model

Write the *effects model* (sum to zero) for this study, make sure to provide the index values, assumptions, and clearly define each term in context.

\vspace{1.8in}

## e. Using R, obtain the parameter/coefficient estimates for the following:

::: callout-caution
## Reminder for R

When using R, you need to change the default settings to use "sum to zero":

```         
options(contrasts = c("contr.sum", "contr.poly"))
mod <- lm(...)
```

Check your factor order with:

```         
levels(data$variable_name)
```
:::

-   $\hat \mu$ = \_\_\_\_\_\_\_\_\_\_\_

-   $\hat \tau_1$ = \_\_\_\_\_\_\_\_\_\_\_

-   $\hat \tau_2$ = \_\_\_\_\_\_\_\_\_\_\_

-   $\hat \tau_3$ = \_\_\_\_\_\_\_\_\_\_\_

for $i=1$ (Pretimed), $i=2$ (Semi-actuated), $i=3$ (Fully actuated).

$$
\begin{align*}
\hat \tau_3 &= -(\hat \tau_1 + \hat \tau_2 ) \\
&= -(12.020 - 2.56) \\
&= -9.452
\end{align*}
$$

$\hat{\mu} = 23.46$

```{r}
emmeans(traffic_mod, ~ traffic_signal)
```

## f. Interpret Parameter Estimates

In context of the data, provide an interpretation for $\hat \tau_1 = 12.02$. It may be helpful to think about what $\bar y_1\cdot$ was from (c).

\vspace{1in}

## g. Provide a conclusion

In context of the data, write out your conclusion from the overall ANOVA F-test from (c). Make sure to provide all evidence -- (F = , df = , p = ).

\vspace{1in}
