---
title: "Activity 3.1: Power Planning for a CRD"
subtitle: "Training Programs & Vertical Jump"
execute: 
  error: true
  message: false
  warning: false

author: Your Name
echo: true
embed-resources: true
format: typst
---
I worked with: _______________________________

```{r}
#| label: setup

# the only package you need today is the pwr2 package

pak::pak("cran/pwr2")
library(pwr2)
```

You are planning a study to evaluate the effect of 5 different training programs (speed, power, agility, endurance, balance) on improvement in vertical jump height (measured in inches).

From previous studies, the experimental error (i.e., $\sigma^2$) is estimated to be about 16 inches.

Talking to a PT, the expected treatment means (in inches) are:

| Program       | Speed | Power | Agility | Endurance | Balance |
|---------------|------:|------:|--------:|----------:|--------:|
| Expected mean |    13 |    11 |      17 |        15 |       9 |

We will use a significance level of $\alpha = 0.05$ and we want 90% power. Even though the configuration above shows "equally spaced" mean configuration, you can use "minimial power" configuration throughout this activity.

## Setup

### a. Identify the study structure:

-   Treatment Structure: (name, factor, levels, treatments)

-   Design Structure: (name, e.u., m.u.; note, we don't know the reps yet!)

-   Response:

### b. What is the value of $\delta$ in this setting? What does this represent (in words)?

## Part 1: Baseline Power

We will treat the $\delta$ from above as our *specific alternative*.

### c. Use R to find the required sample size per group $r$ to achieve 90% power.

::: callout-tip
## What information do you have?

-   t = _______
-   $\hat\sigma^2$ = _______
-   $\delta$ = _______ *Use minimal power configuration!*
-   Set $\alpha =$ _______
-   Desired Power = _______
:::

```{r}
# code to calculate sample size
```

Report your result:

-   Required $r$ (per group) = _______

## Part 2: What changes the required sample size?

Complete the table below. For each scenario, you do **NOT** need to show steps, just use software and record the new required $r$.

::: callout-note
The point here is **direction + magnitude**: how design decisions change the planning.
:::

### d. Fill in the table

Start with the baseline for each scenario and only make the specified changes (i.e., you always go back to the input information you specified in (c))

```{r}
# code to calculate sample sizes for table
```

| Scenario change | What you change | New required $r$ | More / fewer players? |
|-----------------|-----------------|--------------------:|-----------------|
| Baseline | $\sigma^2=16$, $\alpha=0.05$, means = (13,11,17,15,9), target power = 0.90 | \_\_\_\_\_ |  |
| Higher variability | $\hat\sigma^2=36$ (instead of $\hat\sigma^2 = 16$) | \_\_\_\_\_ | More  /  Fewer? |
| More conservative test | $\alpha=0.01$ (instead of 0.05) | \_\_\_\_\_ | More  /  Fewer? |
| Smaller effect | means = (11,10,13,12,9) | \_\_\_\_\_ | More  /  Fewer? |
| Fewer groups | only 3 programs with means (9,13,17) | \_\_\_\_\_ | More  /  Fewer? |

## Part 3: What happens to power if we *can’t* afford the sample size?

### e. Suppose you only have $r = 2$ players per group.

Using the **baseline** setup ($\hat\sigma^2=16$, $\alpha=0.05$, baseline means), compute the resulting power using software.

```{r}
# code to calculate power
```

-   Power when $r=2$ = _______

### f. Interpret the power from (e) in one sentence (in context).

### g. Suppose you only have $r = 2$ players per group, what is the difference you could detect at 90% power?

Using the **baseline** setup ($\hat\sigma^2=16$, $\alpha=0.05$), compute the difference $\delta$.

*If using R, you will have to use a guess & check method.*

```{r}
# code to calculate smallest detectable difference
```

-   Detectable difference at a power of 90% when $r=2$: $\delta$ = _____________

## Part 4: Estimating $\hat\sigma^2$

In practice, estimating the experimental error before collecting data is difficult. Two common approaches are:

1. Use previous literature and domain knowledge.
2. Run a pilot study.

### h. Use [Google Scholar](https://scholar.google.com/) and/or the [Cal Poly Library OneSearch](https://csu-calpoly.primo.exlibrisgroup.com/discovery/search?vid=01CALS_PSU:01CALS_PSU&lang=en) to find one **academic paper** related to the effect of training programs on vertical jump height. 

Provide the following:

+ Your search prompt (keywords):

+ A paper that could be useful for estimating experimental error *(follow the DOI or journal link to the actual paper)*:

+ Brief justification (1-2 sentences): Why might this paper be useful for estimating $\sigma^2$? *(Think: sample size, number of groups, reported SDs, similar response units.)*
