---
title: "Lab 2 Analysis Appendix"
subtitle: "YOUR REPORT TITLE"
author: "YOUR NAMES"
execute: 
  error: true
  message: false
  warning: false
  echo: true
embed-resources: true
format: html
---

```{r}
#| label: setup

# I believe these are the main packages you will need
# add any more that you end up needing here!

library(tidyverse)
library(emmeans)
library(pwr2)
library(multcomp)
```

## Power Analysis

Step 1: determine $\sigma^2$

```{r}
# code here if needed

```

Step 2: calculate sample size to achieve 80% power, using the $\sigma^2$, $\delta$, and $t$ for your study and $\alpha = 0.05$.

```{r}
# code here if needed

```


## Random Sampling & Treatment Assignment

We haven't talked about `set.seed()` yet in class, but it makes random procedures give you the same output every time! This makes your results **reproducible**. You are welcome to change the numbers that I provided before you start, but you shouldn't change them once you have determined your sample and your assignment.

First, randomly sample the required number of participants based on your power analysis ($r\times t$).

```{r}
# code for random sampling
set.seed(243)

N <- ### FILL THIS IN WITH YOUR SAMPLE SIZE

sample_ids <- sample(1:60, size = N, replace = F)

sort(sample_ids)
```

Then, randomly assign treatments to participants using a CRD.

Call the dataset you create `crd_table`

See instructions for additional details.

```{r}
# code for creating random assignment dataset
set.seed(393)

crd_table <- ###.... CODE
  

```

Save your table for data entry:

```{r}
write_csv(crd_table,
          file = "lab2_crd_randomization.csv")
```

## Data Analysis

Read in your study data.

Don't forget to change your treatment variable into a FACTOR variable type!

  - e.g. in Activity 2.2: `mutate(companions = factor(companions))`
  
Check the order of the levels of your variable. Do you want to change the order?

```{r}
# read in data

```

### EDA

Fill in the ____ below in the code to create your plot! Typically, the EV should be the x-axis and the RV should be the y-axis.

```{r}
# code for side-by-side boxplots + individual points

ggplot(data = ___________,
       aes(x = __________,
           y = __________)) +
  geom_boxplot() +
  geom_jitter(alpha = .6, 
              color = "blue",
              height = 0,
              width = .1) +
  labs(title = "GIVE THIS A GOOD TITLE",
       x = "GOOD EV AXIS LABEL",
       y = "GOOD RV AXIS LABEL")

```


### ANOVA


```{r}
# code to conduct ANOVA 

options(contrasts = c("contr.sum", "contr.poly"))


```

### Estimate Model Parameters


```{r}
# code to estimate model parameters


```


### Contrasts or Pairwise Comparisons


```{r}
# code to estimate model parameters


```

### Assess Model Fit


```{r}
# code for residual vs fitted value plot


```

```{r}
# code for qqplot


```