---
title: "Lab 3 Analysis Appendix"
subtitle: "YOUR REPORT TITLE"
author: "YOUR NAMES"
execute: 
  error: true
  message: false
  warning: false
  echo: true
embed-resources: true
format: html
---

```{r}
#| label: setup

# I believe these are the main packages you will need
# add any more that you end up needing here!

library(tidyverse)
library(emmeans)
library(multcomp)
library(multcompView)
library(edibble)
```


## Random Sampling & Treatment Assignment

### Random Sample

First, randomly sample 48 participants

```{r}
# code for random sampling
set.seed(340)

N <- 48

sample_ids <- sample(1:60, size = N, replace = F)

sort(sample_ids)
```

### Random Assignment + Data Table

Then, randomly assign treatments to participants using a CRD.

Call the dataset you create `crd_3factor_table`

See instructions for additional details -- you will want to use the `edibble` package.

```{r}
# code for creating random assignment dataset
set.seed(3334)

# create edibble design object

des <- ### EDIBBLE CODE

# create table WITH subsamples

crd_3factor_table <- serve_table(des) |> 
  expand_grid(subsample = 1:2) |> 
  mutate(subsample = paste0("subsample0", subsample))

# add response variable(s) -- 
# if you are doing blood pressure you might want 
# one column for systolic and one for diastolic

crd_3factor_table <- crd_3factor_table |> 
  mutate( )

```

Save your table to your computer for data entry:

```{r}
write_csv(crd_3factor_table,
          file = "lab3_crd_randomization.csv")
```

You will want to copy and paste the data from that .csv to your shared Google sheet or OneDrive!

## Data Analysis

### Read in Data

Read in your study data.

Don't forget to change your treatment variables into FACTOR variable types!

  - e.g. in Activity 2.2: `mutate(companions = factor(companions))`
  
Check the order of the levels of your variable. Do you want to change the order?

```{r}
# read in data

```

### Clean Data

:::callout-warning
Clean up your systolic / diastolic column if it is not already two columns!!

This code is assuming that you entered your data in a column called `bp` like `"systolic/diastolic"` -- **delete anything you don't need and change variable names as needed!!** Ask Dr. C if you need help with this.
:::

```{r}
# clean up blood pressure variable if needed and calculate MAP

# YOU MAY NEED TO CHANGE / DELETE THINGS

data_clean <- data |> 
  separate(col = bp,
           into = c("systolic", "diastolic"),
           sep = "\\/",
           remove = F, convert = T) |> 
  mutate(map = (systolic + (2*diastolic))/3)

```

Aggregate your data to the islander level

```{r}
# use summarize and mean to get islander-level data


```

### Analysis

Create R chunks as you need them for the analysis!

