---
title: "Activity 4.2: Factorial Analysis"
subtitle: "Pig Feed Experiment"
execute: 
  error: true
  message: false
  warning: false

author: Your Name
echo: true
embed-resources: true
format: typst
---
```{r}
#| label: setup

# load needed packages -- you may need to add

library(tidyverse)
library(janitor)
library(emmeans)
library(multcomp)
```

A study was performed to investigate the effects of two feed additives on the growth of male pigs. The additives were Lysine (at 0%, 0.05%, 0.10% and 0.15% added to feed) and Soy Protein (12% or 14%). There were four pigs assigned at random to each treatment combination and each pig was randomly assigned to a separate pen in the building. The measured response was Average Daily Gain (ADG), in pounds, of each pig, over a 7-day period on the experimental feeds. The data are in the file `pigfeed_data.csv` found on Canvas.

::: callout-warning
Ensure `lysine` and `protein` are a character/factor data types after reading the data into R. By default it will read these in as numeric. Code is provided in the .qmd template to do this!
:::

```{r}
# REMINDER -- YOU WILL NEED TO add your file path to the code below 
# in read_csv()

pig_dat <- read_csv("") |> 
  clean_names() |> 
  mutate(lysine = factor(lysine),
         protein = factor(protein))
  
levels(pig_dat$lysine)
levels(pig_dat$protein)
```

The researchers would like to determine which of these additives, if any, affect the pigs’ weight gain, and what trends there are, if any, among the different levels of the additives. In addition, the researchers would like to know if there is a combination of lysine and soy protein which results in the most weight gain over the 7-day period.

## a. Identify the structure of the experiment

**Treatment structure**

**Experimental structure**

## b. Complete the skeleton ANOVA table below (delete rows you don't need)

| Source of Variation                  |DF        |
|--------------------------------------|----------|
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |
|                                      |          |

## c. State the statistical effects model

Make sure to define **all** terms, denote $i,j,k$ values, and include assumptions.

## d. Fit the factorial model in R

```{r}
# model code
```

```{r}
# check assumptions code 
```

-   Following the decision flow chart, provide conclusions for the *relevant* effects. Make sure to support your answer using (F = ; df = ; p = ).

-   Based on your conclusion above, should you proceed by looking at (circle or only keep one):

**B. Joint / Simple Effects** Since there is evidence of an interaction.

 A. Main Effects                  
 B. Joint/Simple Effects

::: callout-note
## $\epsilon_{ijk} \text{ iid } \sim N(0,\sigma^2)$

Get in the habit of looking at your residual plots for model assumptions here!
:::

## e. Visualizing the interaction

-   Create a plot that illustrates how Lysine (x-axis) modifies the effect of Soy Protein (separate colored lines) on ADG.
-   Copy/paste/sketch your plot (if you are not using this .qmd and integrating your plot directly.)
-   In 2–3 sentences, describe what you notice in the plot.

```{r}
# plot code 
```

## f. Simple effects: Protein within Lysine

Conduct slice tests to compare mean ADG between 12% and 14% Protein at each level of Lysine.

```{r}
# slice test code
```

This aligns with the plot, which showed evidence of an effect of protein amount when lysine is 0 or .05, but not evidence of an effect of protein amount when lysine is .1 or .15 (the intervals overlap).

-   Lysine = 0: F = \_\_\_\_\_\_\_; df = \_\_\_\_\_, \_\_\_\_\_; p = \_\_\_\_\_\_\_\_

-   Lysine = 0.05: F = \_\_\_\_\_\_\_; df = \_\_\_\_\_, \_\_\_\_\_; p = \_\_\_\_\_\_\_\_

-   Lysine = 0.1: F = \_\_\_\_\_\_\_; df = \_\_\_\_\_, \_\_\_\_\_; p = \_\_\_\_\_\_\_\_

-   Lysine = 015: F = \_\_\_\_\_\_\_; df = \_\_\_\_\_, \_\_\_\_\_; p = \_\_\_\_\_\_\_\_

Summarize your findings in words, how does this support what you saw in the interaction plot above?

## g. Recommendation

Which feed combination(s) would you recommend to maximize ADG, on average?

Support your recommendation with:

-   treatment means (with units),

-   and appropriate multiple comparisons (e.g., compact letter display).

```{r}
cld(pig_lsmeans, Letters = LETTERS, 
    decreasing = T, adjust = "tukey")
```
We recommend using 14% soy protein additive and either 0.05% or 0% lycine additive. The estimated mean ADG for male pigs fed 14% protein and 0.05% lycine was 3.12 pounds, which was not significantly different from that for male pigs fed 14% protein and 0% lycine of 3 pounds.
